---
title: "N11423579 SIU Wing Hei Final Report"
output:
  pdf_document:
    latex_engine: xelatex
---
```{r}
train <- read.csv("C:/Users/siusa/Desktop/QUT課程/Y2S2/MXN442_24se2 Modern Statistical Computing Techniques/Assignment 3 Final Report/Rstudio/Open Titanic data/train.csv")
```

```{r}
test <- read.csv("C:/Users/siusa/Desktop/QUT課程/Y2S2/MXN442_24se2 Modern Statistical Computing Techniques/Assignment 3 Final Report/Rstudio/Open Titanic data/test.csv")
```

```{r}
test$Name[415] <- "Oliva y Ocana, Miss. Fermina"# becuase Oliva y Ocana, Dona. Fermina do not have Nickname
```


```{r}
gender_submission <- read.csv("C:/Users/siusa/Desktop/QUT課程/Y2S2/MXN442_24se2 Modern Statistical Computing Techniques/Assignment 3 Final Report/Rstudio/Open Titanic data/gender_submission.csv")
```

# Data processing

```{r}
colnames(train)
```

```{r}
str(train)
```

```{r}
str(test)
```

```{r}
str(gender_submission)
```

```{r}
changedtpye_train<-data.frame(PassengerId = train$PassengerId,
                       Survived = as.factor(train$Survived),
                       Pclass = as.factor(train$Pclass),
                       Name = train$Name,
                       Sex = as.factor(train$Sex),
                       Age = train$Age,
                       SibSp = train$SibSp,
                       Parch = train$Parch,
                       Ticket = train$Ticket,
                       Fare = train$Fare,
                       Cabin = train$Cabin,
                       Embarked = train$Embarked
                       )

str(changedtpye_train)
```

```{r}
changedtpye_test<-data.frame(PassengerId = test$PassengerId,
                       Pclass = as.factor(test$Pclass),
                       Name = test$Name,
                       Sex = as.factor(test$Sex),
                       Age = test$Age,
                       SibSp = test$SibSp,
                       Parch = test$Parch,
                       Ticket = test$Ticket,
                       Fare = test$Fare,
                       Cabin = test$Cabin,
                       Embarked = test$Embarked
                       )

str(changedtpye_test)
```

```{r}
changedtpye_gender_submission<-data.frame(PassengerId = gender_submission$PassengerId,
                       Survived = as.factor(gender_submission$Survived)
                       )

str(changedtpye_gender_submission)
```

```{r}
process_train <- subset(changedtpye_train, select = -PassengerId)

# change age to mean
train_meanofAge <- mean(process_train$Age, na.rm = TRUE)
process_train$Age[is.na(process_train$Age)] <- train_meanofAge


process_train <- subset(process_train, select = -Ticket)

process_train <- subset(process_train, select = -Fare)

process_train <- subset(process_train, select = -Cabin)

process_train <- subset(process_train, select = -Embarked)

process_train$Nickname <- sub(".*,\\s*(\\w+)\\..*", "\\1", process_train$Name)
process_train$Nickname <- as.factor(process_train$Nickname)

process_train <- subset(process_train, select = -Name)#must do this after already change Nickname


process_train
```

```{r}
process_test <- subset(changedtpye_test, select = -PassengerId)

# change age to mean
test_meanofAge <- mean(process_test$Age, na.rm = TRUE)
process_test$Age[is.na(process_test$Age)] <- test_meanofAge


process_test <- subset(process_test, select = -Ticket)

process_test <- subset(process_test, select = -Fare)

process_test <- subset(process_test, select = -Cabin)

process_test <- subset(process_test, select = -Embarked)

process_test$Nickname <- sub(".*,\\s*(\\w+)\\..*", "\\1", process_test$Name)
process_test$Nickname <- as.factor(process_test$Nickname)
#process_test$Nickname[415] <- "Miss" # becuase Oliva y Ocana, Dona. Fermina do not have Nickname

process_test <- subset(process_test, select = -Name)#must do this after already change Nickname


process_test
```




```{r}
library(caret)
library(e1071)
library(randomForest)
library(class)
```


```{r}
sum(is.na(process_train))
```

```{r}
sum(is.na(process_test))
```

```{r}
sum(is.na(changedtpye_gender_submission))

```

# Article Model



# Logistic
## Logistic with out cross-validation

```{r}


process_test$Sex <- factor(process_test$Sex, levels = levels(process_train$Sex))
process_test$Nickname <- factor(process_test$Nickname, levels = levels(process_train$Nickname))
set.seed(1)
logisticwithoutcross <- glm(Survived ~ Pclass + Sex + Age + SibSp + Parch + Nickname, # + Nickname
                      data = process_train, 
                      family = binomial)


summary(logisticwithoutcross)


#pred_logisticwithoutcross <- predict(logisticwithoutcross, newdata = process_test, type = "response")
pred_logisticwithoutcross <- predict(logisticwithoutcross, newdata = process_test)


pred_value_logisticwithoutcross <- ifelse(pred_logisticwithoutcross > 0.5, 1, 0)#, type = "response"
#pred_value_logisticwithoutcross <- pred_logisticwithoutcross


conmatrix_logisticwithoutcross <- table(pred_value_logisticwithoutcross, changedtpye_gender_submission$Survived)
print(conmatrix_logisticwithoutcross)

```

```{r}
TP_logisticwithoutcross <- sum(pred_value_logisticwithoutcross == 1 & changedtpye_gender_submission$Survived == 1)

TN_logisticwithoutcross <- sum(pred_value_logisticwithoutcross == 0 & changedtpye_gender_submission$Survived == 0)

FP_logisticwithoutcross <- sum(pred_value_logisticwithoutcross == 1 & changedtpye_gender_submission$Survived == 0)

FN_logisticwithoutcross <- sum(pred_value_logisticwithoutcross == 0 & changedtpye_gender_submission$Survived == 1)

cat("True Positives in Titanic Logistic (TP):", TP_logisticwithoutcross, "\n")
cat("True Negatives in Titanic Logistic (TN):", TN_logisticwithoutcross, "\n")
cat("False Positives in Titanic Logistic (FP):", FP_logisticwithoutcross, "\n")
cat("False Negatives in Titanic Logistic (FN):", FN_logisticwithoutcross, "\n")

```

```{r}
table(pred_value_logisticwithoutcross, changedtpye_gender_submission$Survived)


acc_logisticwithoutcross <- (TP_logisticwithoutcross + TN_logisticwithoutcross) / (TP_logisticwithoutcross + TN_logisticwithoutcross + FP_logisticwithoutcross + FN_logisticwithoutcross)

precision_logisticwithoutcross <- (TP_logisticwithoutcross) / (TP_logisticwithoutcross + FP_logisticwithoutcross)

reca_logisticwithoutcross <- (TP_logisticwithoutcross) / (TP_logisticwithoutcross + FN_logisticwithoutcross)

f1_logisticwithoutcross <- 2 * ((precision_logisticwithoutcross * reca_logisticwithoutcross) / (precision_logisticwithoutcross + reca_logisticwithoutcross))

cat("Accuracy of Titanic Logistic:", acc_logisticwithoutcross, "\n")
cat("Precision of Titanic Logistic:", precision_logisticwithoutcross, "\n")
cat("Recall of Titanic Logistic:", reca_logisticwithoutcross, "\n")
cat("F1-Score of Titanic Logistic:", f1_logisticwithoutcross, "\n")

```

## Logistic with k-fold cross-validation
### Logistic with k-fold cross-validation (k=10)

```{r}


process_test$Sex <- factor(process_test$Sex, levels = levels(process_train$Sex))
process_test$Nickname <- factor(process_test$Nickname, levels = levels(process_train$Nickname))

set.seed(1)
k10 <- trainControl(method = "cv", number = 10)

logistick10 <- train(Survived ~ Pclass + Sex + Age + SibSp + Parch + Nickname, 
                           data = process_train, 
                           method = "glm", 
                           family = binomial, 
                           trControl = k10)

print(logistick10)

summary(logistick10)


pred_logistick10 <- predict(logistick10, newdata = process_test)#, type = "response"


#pred_value_logistick10 <- ifelse(pred_logistick10 > 0.5, 1, 0)#, type = "response"
pred_value_logistick10 <- pred_logistick10


conmatrix_logistick10 <- table(pred_value_logistick10, changedtpye_gender_submission$Survived)
print(conmatrix_logistick10)

```

```{r}
TP_logistick10 <- sum(pred_value_logistick10 == 1 & changedtpye_gender_submission$Survived == 1)

TN_logistick10 <- sum(pred_value_logistick10 == 0 & changedtpye_gender_submission$Survived == 0)

FP_logistick10 <- sum(pred_value_logistick10 == 1 & changedtpye_gender_submission$Survived == 0)

FN_logistick10 <- sum(pred_value_logistick10 == 0 & changedtpye_gender_submission$Survived == 1)

cat("True Positives in Titanic Logistic (TP):", TP_logistick10, "\n")
cat("True Negatives in Titanic Logistic (TN):", TN_logistick10, "\n")
cat("False Positives in Titanic Logistic (FP):", FP_logistick10, "\n")
cat("False Negatives in Titanic Logistic (FN):", FN_logistick10, "\n")
```

```{r}
table(pred_value_logistick10, changedtpye_gender_submission$Survived)


acc_logistick10 <- (TP_logistick10 + TN_logistick10) / (TP_logistick10 + TN_logistick10 + FP_logistick10 + FN_logistick10)

precision_logistick10 <- (TP_logistick10) / (TP_logistick10 + FP_logistick10)

reca_logistick10 <- (TP_logistick10) / (TP_logistick10 + FN_logistick10)

f1_logistick10 <- 2 * ((precision_logistick10 * reca_logistick10) / (precision_logistick10 + reca_logistick10))

cat("Accuracy of Titanic Logistic:", acc_logistick10, "\n")
cat("Precision of Titanic Logistic:", precision_logistick10, "\n")
cat("Recall of Titanic Logistic:", reca_logistick10, "\n")
cat("F1-Score of Titanic Logistic:", f1_logistick10, "\n")

```

### Logistic with k-fold cross-validation (k=5)

```{r}


process_test$Sex <- factor(process_test$Sex, levels = levels(process_train$Sex))
process_test$Nickname <- factor(process_test$Nickname, levels = levels(process_train$Nickname))

set.seed(1)
k5 <- trainControl(method = "cv", number = 5)

logistick5 <- train(Survived ~ Pclass + Sex + Age + SibSp + Parch + Nickname, 
                           data = process_train, 
                           method = "glm", 
                           family = binomial, 
                           trControl = k5)

print(logistick5)

summary(logistick5)


pred_logistick5 <- predict(logistick5, newdata = process_test)#, type = "response"


#pred_value_logistick5 <- ifelse(pred_logistick5 > 0.5, 1, 0)#, type = "response"
pred_value_logistick5 <- pred_logistick5


conmatrix_logistick5 <- table(pred_value_logistick5, changedtpye_gender_submission$Survived)
print(conmatrix_logistick5)

```

```{r}
TP_logistick5 <- sum(pred_value_logistick5 == 1 & changedtpye_gender_submission$Survived == 1)

TN_logistick5 <- sum(pred_value_logistick5 == 0 & changedtpye_gender_submission$Survived == 0)

FP_logistick5 <- sum(pred_value_logistick5 == 1 & changedtpye_gender_submission$Survived == 0)

FN_logistick5 <- sum(pred_value_logistick5 == 0 & changedtpye_gender_submission$Survived == 1)

cat("True Positives in Titanic Logistic (TP):", TP_logistick5, "\n")
cat("True Negatives in Titanic Logistic (TN):", TN_logistick5, "\n")
cat("False Positives in Titanic Logistic (FP):", FP_logistick5, "\n")
cat("False Negatives in Titanic Logistic (FN):", FN_logistick5, "\n")

```

```{r}
table(pred_value_logistick5, changedtpye_gender_submission$Survived)


acc_logistick5 <- (TP_logistick5 + TN_logistick5) / (TP_logistick5 + TN_logistick5 + FP_logistick5 + FN_logistick5)

precision_logistick5 <- (TP_logistick5) / (TP_logistick5 + FP_logistick5)

reca_logistick5 <- (TP_logistick5) / (TP_logistick5 + FN_logistick5)

f1_logistick5 <- 2 * ((precision_logistick5 * reca_logistick5) / (precision_logistick5 + reca_logistick5))

cat("Accuracy of Titanic Logistic:", acc_logistick5, "\n")
cat("Precision of Titanic Logistic:", precision_logistick5, "\n")
cat("Recall of Titanic Logistic:", reca_logistick5, "\n")
cat("F1-Score of Titanic Logistic:", f1_logistick5, "\n")

```

### Logistic with k-fold cross-validation (k=4)

```{r}


process_test$Sex <- factor(process_test$Sex, levels = levels(process_train$Sex))
process_test$Nickname <- factor(process_test$Nickname, levels = levels(process_train$Nickname))

set.seed(1)
k4 <- trainControl(method = "cv", number = 4)

logistick4 <- train(Survived ~ Pclass + Sex + Age + SibSp + Parch + Nickname, 
                           data = process_train, 
                           method = "glm", 
                           family = binomial, 
                           trControl = k4)

print(logistick4)

summary(logistick4)


pred_logistick4 <- predict(logistick4, newdata = process_test)#, type = "response"


#pred_value_logistick4 <- ifelse(pred_logistick4 > 0.5, 1, 0)#, type = "response"
pred_value_logistick4 <- pred_logistick4


conmatrix_logistick4 <- table(pred_value_logistick4, changedtpye_gender_submission$Survived)
print(conmatrix_logistick4)

```

```{r}
TP_logistick4 <- sum(pred_value_logistick4 == 1 & changedtpye_gender_submission$Survived == 1)

TN_logistick4 <- sum(pred_value_logistick4 == 0 & changedtpye_gender_submission$Survived == 0)

FP_logistick4 <- sum(pred_value_logistick4 == 1 & changedtpye_gender_submission$Survived == 0)

FN_logistick4 <- sum(pred_value_logistick4 == 0 & changedtpye_gender_submission$Survived == 1)

cat("True Positives in Titanic Logistic (TP):", TP_logistick4, "\n")
cat("True Negatives in Titanic Logistic (TN):", TN_logistick4, "\n")
cat("False Positives in Titanic Logistic (FP):", FP_logistick4, "\n")
cat("False Negatives in Titanic Logistic (FN):", FN_logistick4, "\n")

```

```{r}
table(pred_value_logistick4, changedtpye_gender_submission$Survived)


acc_logistick4 <- (TP_logistick4 + TN_logistick4) / (TP_logistick4 + TN_logistick4 + FP_logistick4 + FN_logistick4)

precision_logistick4 <- (TP_logistick4) / (TP_logistick4 + FP_logistick4)

reca_logistick4 <- (TP_logistick4) / (TP_logistick4 + FN_logistick4)

f1_logistick4 <- 2 * ((precision_logistick4 * reca_logistick4) / (precision_logistick4 + reca_logistick4))

cat("Accuracy of Titanic Logistic:", acc_logistick4, "\n")
cat("Precision of Titanic Logistic:", precision_logistick4, "\n")
cat("Recall of Titanic Logistic:", reca_logistick4, "\n")
cat("F1-Score of Titanic Logistic:", f1_logistick4, "\n")

```

## Result is the same, one of the reason is data set is to small.



# K nearest neighbors(KNN)
## KNN with out cross-validation

```{r}
library(class)    
library(caret)
#install.packages("kknn")
library(kknn)
```

```{r}

library(kknn)  


process_test$Sex <- factor(process_test$Sex, levels = levels(process_train$Sex))
process_test$Nickname <- factor(process_test$Nickname, levels = levels(process_train$Nickname))



set.seed(1)  
KNNwithoutcross <- kknn(Survived ~ Pclass + Sex + Age + SibSp + Parch + Nickname, train = process_train, test = process_test, k = 10)


pre_KNNwithoutcross <- predict(KNNwithoutcross, newdata = process_test)


conmatrix_KNNwithoutcross <- table(pre_KNNwithoutcross , changedtpye_gender_submission$Survived)


print(conmatrix_KNNwithoutcross)
```

```{r}
TP_KNNwithoutcross <- sum(pre_KNNwithoutcross == 1 & changedtpye_gender_submission$Survived == 1)

TN_KNNwithoutcross <- sum(pre_KNNwithoutcross == 0 & changedtpye_gender_submission$Survived == 0)

FP_KNNwithoutcross <- sum(pre_KNNwithoutcross == 1 & changedtpye_gender_submission$Survived == 0)

FN_KNNwithoutcross <- sum(pre_KNNwithoutcross == 0 & changedtpye_gender_submission$Survived == 1)

cat("True Positives in Titanic KNN:", TP_KNNwithoutcross, "\n")
cat("True Negatives in Titanic KNN:", TN_KNNwithoutcross, "\n")
cat("False Positives in Titanic KNN:", FP_KNNwithoutcross, "\n")
cat("False Negatives in Titanic KNN:", FN_KNNwithoutcross, "\n")

```

```{r}
table(pre_KNNwithoutcross, changedtpye_gender_submission$Survived)


acc_KNNwithoutcross <- (TP_KNNwithoutcross + TN_KNNwithoutcross) / (TP_KNNwithoutcross + TN_KNNwithoutcross + FP_KNNwithoutcross + FN_KNNwithoutcross)

precision_KNNwithoutcross <- (TP_KNNwithoutcross) / (TP_KNNwithoutcross + FP_KNNwithoutcross)

reca_KNNwithoutcross <- (TP_KNNwithoutcross) / (TP_KNNwithoutcross + FN_KNNwithoutcross)

f1_KNNwithoutcross <- 2 * ((precision_KNNwithoutcross * reca_KNNwithoutcross) / (precision_KNNwithoutcross + reca_KNNwithoutcross))

cat("Accuracy of Titanic KNN:", acc_KNNwithoutcross, "\n")
cat("Precision of Titanic KNN:", precision_KNNwithoutcross, "\n")
cat("Recall of Titanic KNN:", reca_KNNwithoutcross, "\n")
cat("F1-Score of Titanic KNN:", f1_KNNwithoutcross, "\n")
```
### when k is bigger in this case, performance is better, I have try 3,5,10,15, 



## KNN with k-fold cross-validation (k=10)
```{r}
library(kknn)  

process_test$Sex <- factor(process_test$Sex, levels = levels(process_train$Sex))
process_test$Nickname <- factor(process_test$Nickname, levels = levels(process_train$Nickname))

set.seed(1)  
k10 <- trainControl(method = "cv", number = 10)

KNNk10 <- train(Survived ~ Pclass + Sex + Age + SibSp + Parch + Nickname, 
                         data = process_train, 
                         method = "kknn", 
                         trControl = k10,
                         tuneLength = 10) #range of 10 different potential k-values to find the optimal one
print(KNNk10)


pre_KNNk10 <- predict(KNNk10, newdata = process_test)


conmatrix_KNNk10 <- table(pre_KNNk10, changedtpye_gender_submission$Survived)


print(conmatrix_KNNk10)
```

```{r}
TP_KNNk10 <- sum(pre_KNNk10 == 1 & changedtpye_gender_submission$Survived == 1)

TN_KNNk10 <- sum(pre_KNNk10 == 0 & changedtpye_gender_submission$Survived == 0)

FP_KNNk10 <- sum(pre_KNNk10 == 1 & changedtpye_gender_submission$Survived == 0)

FN_KNNk10 <- sum(pre_KNNk10 == 0 & changedtpye_gender_submission$Survived == 1)

cat("True Positives in Titanic KNN:", TP_KNNk10, "\n")
cat("True Negatives in Titanic KNN:", TN_KNNk10, "\n")
cat("False Positives in Titanic KNN:", FP_KNNk10, "\n")
cat("False Negatives in Titanic KNN:", FN_KNNk10, "\n")

```

```{r}
table(pre_KNNk10, changedtpye_gender_submission$Survived)


acc_KNNk10 <- (TP_KNNk10 + TN_KNNk10) / (TP_KNNk10 + TN_KNNk10 + FP_KNNk10 + FN_KNNk10)

precision_KNNk10 <- (TP_KNNk10) / (TP_KNNk10 + FP_KNNk10)

reca_KNNk10 <- (TP_KNNk10) / (TP_KNNk10 + FN_KNNk10)

f1_KNNk10 <- 2 * ((precision_KNNk10 * reca_KNNk10) / (precision_KNNk10 + reca_KNNk10))

cat("Accuracy of Titanic KNN:", acc_KNNk10, "\n")
cat("Precision of Titanic KNN:", precision_KNNk10, "\n")
cat("Recall of Titanic KNN:", reca_KNNk10, "\n")
cat("F1-Score of Titanic KNN:", f1_KNNk10, "\n")
```


## KNN with k-fold cross-validation (k=5)
```{r}
library(kknn)  

process_test$Sex <- factor(process_test$Sex, levels = levels(process_train$Sex))
process_test$Nickname <- factor(process_test$Nickname, levels = levels(process_train$Nickname))

set.seed(1)  
k5 <- trainControl(method = "cv", number = 5)

KNNk5 <- train(Survived ~ Pclass + Sex + Age + SibSp + Parch + Nickname, 
                         data = process_train, 
                         method = "kknn", 
                         trControl = k5,
                         tuneLength = 10) #range of 10 different potential k-values to find the optimal one
print(KNNk5)


pre_KNNk5 <- predict(KNNk5, newdata = process_test)


conmatrix_KNNk5 <- table(pre_KNNk5, changedtpye_gender_submission$Survived)


print(conmatrix_KNNk5)
```

```{r}
TP_KNNk5 <- sum(pre_KNNk5 == 1 & changedtpye_gender_submission$Survived == 1)

TN_KNNk5 <- sum(pre_KNNk5 == 0 & changedtpye_gender_submission$Survived == 0)

FP_KNNk5 <- sum(pre_KNNk5 == 1 & changedtpye_gender_submission$Survived == 0)

FN_KNNk5 <- sum(pre_KNNk5 == 0 & changedtpye_gender_submission$Survived == 1)

cat("True Positives in Titanic KNN:", TP_KNNk5, "\n")
cat("True Negatives in Titanic KNN:", TN_KNNk5, "\n")
cat("False Positives in Titanic KNN:", FP_KNNk5, "\n")
cat("False Negatives in Titanic KNN:", FN_KNNk5, "\n")
```

```{r}
table(pre_KNNk5, changedtpye_gender_submission$Survived)


acc_KNNk5 <- (TP_KNNk5 + TN_KNNk5) / (TP_KNNk5 + TN_KNNk5 + FP_KNNk5 + FN_KNNk5)

precision_KNNk5 <- (TP_KNNk5) / (TP_KNNk5 + FP_KNNk5)

reca_KNNk5 <- (TP_KNNk5) / (TP_KNNk5 + FN_KNNk5)

f1_KNNk5 <- 2 * ((precision_KNNk5 * reca_KNNk5) / (precision_KNNk5 + reca_KNNk5))

cat("Accuracy of Titanic KNN:", acc_KNNk5, "\n")
cat("Precision of Titanic KNN:", precision_KNNk5, "\n")
cat("Recall of Titanic KNN:", reca_KNNk5, "\n")
cat("F1-Score of Titanic KNN:", f1_KNNk5, "\n")
```

## KNN with k-fold cross-validation (k=4)
```{r}
library(kknn)  

process_test$Sex <- factor(process_test$Sex, levels = levels(process_train$Sex))
process_test$Nickname <- factor(process_test$Nickname, levels = levels(process_train$Nickname))

set.seed(1)  
k4 <- trainControl(method = "cv", number = 4)

KNNk4 <- train(Survived ~ Pclass + Sex + Age + SibSp + Parch + Nickname, 
                         data = process_train, 
                         method = "kknn", 
                         trControl = k4,
                         tuneLength = 10) #range of 10 different potential k-values to find the optimal one
print(KNNk4)


pre_KNNk4 <- predict(KNNk4, newdata = process_test)


conmatrix_KNNk4 <- table(pre_KNNk4, changedtpye_gender_submission$Survived)


print(conmatrix_KNNk4)
```

```{r}
TP_KNNk4 <- sum(pre_KNNk4 == 1 & changedtpye_gender_submission$Survived == 1)

TN_KNNk4 <- sum(pre_KNNk4 == 0 & changedtpye_gender_submission$Survived == 0)

FP_KNNk4 <- sum(pre_KNNk4 == 1 & changedtpye_gender_submission$Survived == 0)

FN_KNNk4 <- sum(pre_KNNk4 == 0 & changedtpye_gender_submission$Survived == 1)

cat("True Positives in Titanic KNN:", TP_KNNk4, "\n")
cat("True Negatives in Titanic KNN:", TN_KNNk4, "\n")
cat("False Positives in Titanic KNN:", FP_KNNk4, "\n")
cat("False Negatives in Titanic KNN:", FN_KNNk4, "\n")
```

```{r}
table(pre_KNNk4, changedtpye_gender_submission$Survived)


acc_KNNk4 <- (TP_KNNk4 + TN_KNNk4) / (TP_KNNk4 + TN_KNNk4 + FP_KNNk4 + FN_KNNk4)

precision_KNNk4 <- (TP_KNNk4) / (TP_KNNk4 + FP_KNNk4)

reca_KNNk4 <- (TP_KNNk4) / (TP_KNNk4 + FN_KNNk4)

f1_KNNk4 <- 2 * ((precision_KNNk4 * reca_KNNk4) / (precision_KNNk4 + reca_KNNk4))

cat("Accuracy of Titanic KNN:", acc_KNNk4, "\n")
cat("Precision of Titanic KNN:", precision_KNNk4, "\n")
cat("Recall of Titanic KNN:", reca_KNNk4, "\n")
cat("F1-Score of Titanic KNN:", f1_KNNk4, "\n")
```

### not the same with Article(kfold higher performance higher), in here k=5 is higher, k=4 is the lower.



# Support Vector Machine(SVM) - radial
## SVM with out cross-validation - radial - Controls variance by squashing down most dimensions(in book)

```{r}
library(e1071)


process_test$Sex <- factor(process_test$Sex, levels = levels(process_train$Sex))
process_test$Nickname <- factor(process_test$Nickname, levels = levels(process_train$Nickname))


set.seed(1)
SVMwithoutcross <- svm(Survived ~ Pclass + Sex + Age + SibSp + Parch + Nickname, 
                       data = process_train, 
                       kernel = "radial")
summary(SVMwithoutcross)

pred_SVMwithoutcross <- predict(SVMwithoutcross, newdata = process_test)


conmatrix_SVMwithoutcross <- table(pred_SVMwithoutcross, changedtpye_gender_submission$Survived)


print(conmatrix_SVMwithoutcross)
```

```{r}
TP_SVMwithoutcross <- sum(pred_SVMwithoutcross == 1 & changedtpye_gender_submission$Survived == 1)

TN_SVMwithoutcross <- sum(pred_SVMwithoutcross == 0 & changedtpye_gender_submission$Survived == 0)

FP_SVMwithoutcross <- sum(pred_SVMwithoutcross == 1 & changedtpye_gender_submission$Survived == 0)

FN_SVMwithoutcross <- sum(pred_SVMwithoutcross == 0 & changedtpye_gender_submission$Survived == 1)

cat("True Positives in Titanic SVM:", TP_SVMwithoutcross, "\n")
cat("True Negatives in Titanic SVM:", TN_SVMwithoutcross, "\n")
cat("False Positives in Titanic SVM:", FP_SVMwithoutcross, "\n")
cat("False Negatives in Titanic SVM:", FN_SVMwithoutcross, "\n")
```

```{r}
table(pred_SVMwithoutcross, changedtpye_gender_submission$Survived)


acc_SVMwithoutcross <- (TP_SVMwithoutcross + TN_SVMwithoutcross) / (TP_SVMwithoutcross + TN_SVMwithoutcross + FP_SVMwithoutcross + FN_SVMwithoutcross)

precision_SVMwithoutcross <- (TP_SVMwithoutcross) / (TP_SVMwithoutcross + FP_SVMwithoutcross)

reca_SVMwithoutcross <- (TP_SVMwithoutcross) / (TP_SVMwithoutcross + FN_SVMwithoutcross)

f1_SVMwithoutcross <- 2 * ((precision_SVMwithoutcross * reca_SVMwithoutcross) / (precision_SVMwithoutcross + reca_SVMwithoutcross))

cat("Accuracy of Titanic SVM:", acc_SVMwithoutcross, "\n")
cat("Precision of Titanic SVM:", precision_SVMwithoutcross, "\n")
cat("Recall of Titanic SVM:", reca_SVMwithoutcross, "\n")
cat("F1-Score of Titanic SVM:", f1_SVMwithoutcross, "\n")
```


## SVM with k-fold cross-validation (k=10) - radial

```{r}
library(e1071)
library(caret)

process_test$Sex <- factor(process_test$Sex, levels = levels(process_train$Sex))
process_test$Nickname <- factor(process_test$Nickname, levels = levels(process_train$Nickname))


set.seed(1)
k10 <- trainControl(method = "cv", number = 10)
SVMk10 <- train(Survived ~ Pclass + Sex + Age + SibSp + Parch + Nickname, 
                data = process_train, 
                method = "svmRadial", 
                trControl = k10,        
                tuneLength = 10)        

print(SVMk10)
summary(SVMk10)

pred_SVMk10 <- predict(SVMk10, newdata = process_test)

conmatrix_SVMk10 <- table(pred_SVMk10, changedtpye_gender_submission$Survived)
print(conmatrix_SVMk10)
```

```{r}
TP_SVMk10 <- sum(pred_SVMk10 == 1 & changedtpye_gender_submission$Survived == 1)

TN_SVMk10 <- sum(pred_SVMk10 == 0 & changedtpye_gender_submission$Survived == 0)

FP_SVMk10 <- sum(pred_SVMk10 == 1 & changedtpye_gender_submission$Survived == 0)

FN_SVMk10 <- sum(pred_SVMk10 == 0 & changedtpye_gender_submission$Survived == 1)

cat("True Positives in Titanic SVM:", TP_SVMk10, "\n")
cat("True Negatives in Titanic SVM:", TN_SVMk10, "\n")
cat("False Positives in Titanic SVM:", FP_SVMk10, "\n")
cat("False Negatives in Titanic SVM:", FN_SVMk10, "\n")
```

```{r}
table(pred_SVMk10, changedtpye_gender_submission$Survived)


acc_SVMk10 <- (TP_SVMk10 + TN_SVMk10) / (TP_SVMk10 + TN_SVMk10 + FP_SVMk10 + FN_SVMk10)

precision_SVMk10 <- (TP_SVMk10) / (TP_SVMk10 + FP_SVMk10)

reca_SVMk10 <- (TP_SVMk10) / (TP_SVMk10 + FN_SVMk10)

f1_SVMk10 <- 2 * ((precision_SVMk10 * reca_SVMk10) / (precision_SVMk10 + reca_SVMk10))

cat("Accuracy of Titanic SVM:", acc_SVMk10, "\n")
cat("Precision of Titanic SVM:", precision_SVMk10, "\n")
cat("Recall of Titanic SVM:", reca_SVMk10, "\n")
cat("F1-Score of Titanic SVM:", f1_SVMk10, "\n")
```

## SVM with k-fold cross-validation (k=5) - radial

```{r}
library(e1071)
library(caret)

process_test$Sex <- factor(process_test$Sex, levels = levels(process_train$Sex))
process_test$Nickname <- factor(process_test$Nickname, levels = levels(process_train$Nickname))


set.seed(1)
k5 <- trainControl(method = "cv", number = 5)
SVMk5 <- train(Survived ~ Pclass + Sex + Age + SibSp + Parch + Nickname, 
                data = process_train, 
                method = "svmRadial",   
                trControl = k5,      
                tuneLength = 10)     

print(SVMk5)
summary(SVMk5)

pred_SVMk5 <- predict(SVMk5, newdata = process_test)

conmatrix_SVMk5 <- table(pred_SVMk5, changedtpye_gender_submission$Survived)
print(conmatrix_SVMk5)
```

```{r}
TP_SVMk5 <- sum(pred_SVMk5 == 1 & changedtpye_gender_submission$Survived == 1)

TN_SVMk5 <- sum(pred_SVMk5 == 0 & changedtpye_gender_submission$Survived == 0)

FP_SVMk5 <- sum(pred_SVMk5 == 1 & changedtpye_gender_submission$Survived == 0)

FN_SVMk5 <- sum(pred_SVMk5 == 0 & changedtpye_gender_submission$Survived == 1)

cat("True Positives in Titanic SVM:", TP_SVMk5, "\n")
cat("True Negatives in Titanic SVM:", TN_SVMk5, "\n")
cat("False Positives in Titanic SVM:", FP_SVMk5, "\n")
cat("False Negatives in Titanic SVM:", FN_SVMk5, "\n")
```

```{r}
table(pred_SVMk5, changedtpye_gender_submission$Survived)


acc_SVMk5 <- (TP_SVMk5 + TN_SVMk5) / (TP_SVMk5 + TN_SVMk5 + FP_SVMk5 + FN_SVMk5)

precision_SVMk5 <- (TP_SVMk5) / (TP_SVMk5 + FP_SVMk5)

reca_SVMk5 <- (TP_SVMk5) / (TP_SVMk5 + FN_SVMk5)

f1_SVMk5 <- 2 * ((precision_SVMk5 * reca_SVMk5) / (precision_SVMk5 + reca_SVMk5))

cat("Accuracy of Titanic SVM", acc_SVMk5, "\n")
cat("Precision of Titanic SVM:", precision_SVMk5, "\n")
cat("Recall of Titanic SVM:", reca_SVMk5, "\n")
cat("F1-Score of Titanic SVM:", f1_SVMk5, "\n")
```

## SVM with k-fold cross-validation (k=4) - radial

```{r}
library(e1071)
library(caret)

process_test$Sex <- factor(process_test$Sex, levels = levels(process_train$Sex))
process_test$Nickname <- factor(process_test$Nickname, levels = levels(process_train$Nickname))


set.seed(1)
k4 <- trainControl(method = "cv", number = 4)
SVMk4 <- train(Survived ~ Pclass + Sex + Age + SibSp + Parch + Nickname, 
                data = process_train, 
                method = "svmRadial",  
                trControl = k4,        
                tuneLength = 10)        

print(SVMk4)
summary(SVMk4)

pred_SVMk4 <- predict(SVMk4, newdata = process_test)

conmatrix_SVMk4 <- table(pred_SVMk4, changedtpye_gender_submission$Survived)
print(conmatrix_SVMk4)
```

```{r}
TP_SVMk4 <- sum(pred_SVMk4 == 1 & changedtpye_gender_submission$Survived == 1)

TN_SVMk4 <- sum(pred_SVMk4 == 0 & changedtpye_gender_submission$Survived == 0)

FP_SVMk4 <- sum(pred_SVMk4 == 1 & changedtpye_gender_submission$Survived == 0)

FN_SVMk4 <- sum(pred_SVMk4 == 0 & changedtpye_gender_submission$Survived == 1)

cat("True Positives in Titanic SVM:", TP_SVMk4, "\n")
cat("True Negatives in Titanic SVM:", TN_SVMk4, "\n")
cat("False Positives in Titanic SVM:", FP_SVMk4, "\n")
cat("False Negatives in Titanic SVM:", FN_SVMk4, "\n")
```

```{r}
table(pred_SVMk4, changedtpye_gender_submission$Survived)


acc_SVMk4 <- (TP_SVMk4 + TN_SVMk4) / (TP_SVMk4 + TN_SVMk4 + FP_SVMk4 + FN_SVMk4)

precision_SVMk4 <- (TP_SVMk4) / (TP_SVMk4 + FP_SVMk4)

reca_SVMk4 <- (TP_SVMk4) / (TP_SVMk4 + FN_SVMk4)

f1_SVMk4 <- 2 * ((precision_SVMk4 * reca_SVMk4) / (precision_SVMk4 + reca_SVMk4))

cat("Accuracy of Titanic SVM:", acc_SVMk4, "\n")
cat("Precision of Titanic SVM:", precision_SVMk4, "\n")
cat("Recall of Titanic SVM:", reca_SVMk4, "\n")
cat("F1-Score of Titanic SVM:", f1_SVMk4, "\n")
```

# Support Vector Machine(SVM) - polynomial
## SVM with out cross-validation - polynomial - Use kernels to introduce nonlinearities in a controlled manner(in book)

```{r}
library(e1071)


process_test$Sex <- factor(process_test$Sex, levels = levels(process_train$Sex))
process_test$Nickname <- factor(process_test$Nickname, levels = levels(process_train$Nickname))


set.seed(1)
SVMwithoutcross <- svm(Survived ~ Pclass + Sex + Age + SibSp + Parch + Nickname, 
                       data = process_train, 
                       kernel = "polynomial", #Nonlinear data and it is polynomial relationships
                       degree = 3)
summary(SVMwithoutcross)

pred_SVMwithoutcross <- predict(SVMwithoutcross, newdata = process_test)


conmatrix_SVMwithoutcross <- table(pred_SVMwithoutcross, changedtpye_gender_submission$Survived)


print(conmatrix_SVMwithoutcross)
```

```{r}
TP_SVMwithoutcross <- sum(pred_SVMwithoutcross == 1 & changedtpye_gender_submission$Survived == 1)

TN_SVMwithoutcross <- sum(pred_SVMwithoutcross == 0 & changedtpye_gender_submission$Survived == 0)

FP_SVMwithoutcross <- sum(pred_SVMwithoutcross == 1 & changedtpye_gender_submission$Survived == 0)

FN_SVMwithoutcross <- sum(pred_SVMwithoutcross == 0 & changedtpye_gender_submission$Survived == 1)

cat("True Positives in Titanic SVM:", TP_SVMwithoutcross, "\n")
cat("True Negatives in Titanic SVM:", TN_SVMwithoutcross, "\n")
cat("False Positives in Titanic SVM:", FP_SVMwithoutcross, "\n")
cat("False Negatives in Titanic SVM:", FN_SVMwithoutcross, "\n")
```

```{r}
table(pred_SVMwithoutcross, changedtpye_gender_submission$Survived)


acc_SVMwithoutcross <- (TP_SVMwithoutcross + TN_SVMwithoutcross) / (TP_SVMwithoutcross + TN_SVMwithoutcross + FP_SVMwithoutcross + FN_SVMwithoutcross)

precision_SVMwithoutcross <- (TP_SVMwithoutcross) / (TP_SVMwithoutcross + FP_SVMwithoutcross)

reca_SVMwithoutcross <- (TP_SVMwithoutcross) / (TP_SVMwithoutcross + FN_SVMwithoutcross)

f1_SVMwithoutcross <- 2 * ((precision_SVMwithoutcross * reca_SVMwithoutcross) / (precision_SVMwithoutcross + reca_SVMwithoutcross))

cat("Accuracy of Titanic SVM:", acc_SVMwithoutcross, "\n")
cat("Precision of Titanic SVM:", precision_SVMwithoutcross, "\n")
cat("Recall of Titanic SVM:", reca_SVMwithoutcross, "\n")
cat("F1-Score of Titanic SVM:", f1_SVMwithoutcross, "\n")
```


## SVM with k-fold cross-validation (k=10) - polynomial

```{r}
library(e1071)
library(caret)

process_test$Sex <- factor(process_test$Sex, levels = levels(process_train$Sex))
process_test$Nickname <- factor(process_test$Nickname, levels = levels(process_train$Nickname))


set.seed(1)
k10 <- trainControl(method = "cv", number = 10)
SVMk10 <- train(Survived ~ Pclass + Sex + Age + SibSp + Parch + Nickname, 
                data = process_train, 
                method = "svmPoly", #Nonlinear data and it is polynomial relationships
                degree = 3, #default by svmPoly
                trControl = k10)     #tuneLength = 10


print(SVMk10)
summary(SVMk10)

pred_SVMk10 <- predict(SVMk10, newdata = process_test)

conmatrix_SVMk10 <- table(pred_SVMk10, changedtpye_gender_submission$Survived)
print(conmatrix_SVMk10)
```

```{r}
TP_SVMk10 <- sum(pred_SVMk10 == 1 & changedtpye_gender_submission$Survived == 1)

TN_SVMk10 <- sum(pred_SVMk10 == 0 & changedtpye_gender_submission$Survived == 0)

FP_SVMk10 <- sum(pred_SVMk10 == 1 & changedtpye_gender_submission$Survived == 0)

FN_SVMk10 <- sum(pred_SVMk10 == 0 & changedtpye_gender_submission$Survived == 1)

cat("True Positives in Titanic SVM:", TP_SVMk10, "\n")
cat("True Negatives in Titanic SVM:", TN_SVMk10, "\n")
cat("False Positives in Titanic SVM:", FP_SVMk10, "\n")
cat("False Negatives in Titanic SVM:", FN_SVMk10, "\n")
```

```{r}
table(pred_SVMk10, changedtpye_gender_submission$Survived)


acc_SVMk10 <- (TP_SVMk10 + TN_SVMk10) / (TP_SVMk10 + TN_SVMk10 + FP_SVMk10 + FN_SVMk10)

precision_SVMk10 <- (TP_SVMk10) / (TP_SVMk10 + FP_SVMk10)

reca_SVMk10 <- (TP_SVMk10) / (TP_SVMk10 + FN_SVMk10)

f1_SVMk10 <- 2 * ((precision_SVMk10 * reca_SVMk10) / (precision_SVMk10 + reca_SVMk10))

cat("Accuracy of Titanic SVM:", acc_SVMk10, "\n")
cat("Precision of Titanic SVM:", precision_SVMk10, "\n")
cat("Recall of Titanic SVM:", reca_SVMk10, "\n")
cat("F1-Score of Titanic SVM:", f1_SVMk10, "\n")
```

## SVM with k-fold cross-validation (k=5) - polynomial

```{r}
library(e1071)
library(caret)

process_test$Sex <- factor(process_test$Sex, levels = levels(process_train$Sex))
process_test$Nickname <- factor(process_test$Nickname, levels = levels(process_train$Nickname))


set.seed(1)
k5 <- trainControl(method = "cv", number = 5)
SVMk5 <- train(Survived ~ Pclass + Sex + Age + SibSp + Parch + Nickname, 
                data = process_train, 
                method = "svmPoly",#Nonlinear data and it is polynomial relationships
                degree = 3, 
                trControl = k5)     

print(SVMk5)
summary(SVMk5)

pred_SVMk5 <- predict(SVMk5, newdata = process_test)

conmatrix_SVMk5 <- table(pred_SVMk5, changedtpye_gender_submission$Survived)
print(conmatrix_SVMk5)
```

```{r}
TP_SVMk5 <- sum(pred_SVMk5 == 1 & changedtpye_gender_submission$Survived == 1)

TN_SVMk5 <- sum(pred_SVMk5 == 0 & changedtpye_gender_submission$Survived == 0)

FP_SVMk5 <- sum(pred_SVMk5 == 1 & changedtpye_gender_submission$Survived == 0)

FN_SVMk5 <- sum(pred_SVMk5 == 0 & changedtpye_gender_submission$Survived == 1)

cat("True Positives in Titanic SVM:", TP_SVMk5, "\n")
cat("True Negatives in Titanic SVM:", TN_SVMk5, "\n")
cat("False Positives in Titanic SVM:", FP_SVMk5, "\n")
cat("False Negatives in Titanic SVM:", FN_SVMk5, "\n")
```

```{r}
table(pred_SVMk5, changedtpye_gender_submission$Survived)


acc_SVMk5 <- (TP_SVMk5 + TN_SVMk5) / (TP_SVMk5 + TN_SVMk5 + FP_SVMk5 + FN_SVMk5)

precision_SVMk5 <- (TP_SVMk5) / (TP_SVMk5 + FP_SVMk5)

reca_SVMk5 <- (TP_SVMk5) / (TP_SVMk5 + FN_SVMk5)

f1_SVMk5 <- 2 * ((precision_SVMk5 * reca_SVMk5) / (precision_SVMk5 + reca_SVMk5))

cat("Accuracy of Titanic SVM", acc_SVMk5, "\n")
cat("Precision of Titanic SVM:", precision_SVMk5, "\n")
cat("Recall of Titanic SVM:", reca_SVMk5, "\n")
cat("F1-Score of Titanic SVM:", f1_SVMk5, "\n")
```

## SVM with k-fold cross-validation (k=4) - polynomial

```{r}
library(e1071)
library(caret)

process_test$Sex <- factor(process_test$Sex, levels = levels(process_train$Sex))
process_test$Nickname <- factor(process_test$Nickname, levels = levels(process_train$Nickname))


set.seed(1)
k4 <- trainControl(method = "cv", number = 4)
SVMk4 <- train(Survived ~ Pclass + Sex + Age + SibSp + Parch + Nickname, 
                data = process_train, 
                method = "svmPoly", #Nonlinear data and it is polynomial relationships
                degree = 3, 
                trControl = k4)        

print(SVMk4)
summary(SVMk4)

pred_SVMk4 <- predict(SVMk4, newdata = process_test)

conmatrix_SVMk4 <- table(pred_SVMk4, changedtpye_gender_submission$Survived)
print(conmatrix_SVMk4)
```

```{r}
TP_SVMk4 <- sum(pred_SVMk4 == 1 & changedtpye_gender_submission$Survived == 1)

TN_SVMk4 <- sum(pred_SVMk4 == 0 & changedtpye_gender_submission$Survived == 0)

FP_SVMk4 <- sum(pred_SVMk4 == 1 & changedtpye_gender_submission$Survived == 0)

FN_SVMk4 <- sum(pred_SVMk4 == 0 & changedtpye_gender_submission$Survived == 1)

cat("True Positives in Titanic SVM:", TP_SVMk4, "\n")
cat("True Negatives in Titanic SVM:", TN_SVMk4, "\n")
cat("False Positives in Titanic SVM:", FP_SVMk4, "\n")
cat("False Negatives in Titanic SVM:", FN_SVMk4, "\n")
```

```{r}
table(pred_SVMk4, changedtpye_gender_submission$Survived)


acc_SVMk4 <- (TP_SVMk4 + TN_SVMk4) / (TP_SVMk4 + TN_SVMk4 + FP_SVMk4 + FN_SVMk4)

precision_SVMk4 <- (TP_SVMk4) / (TP_SVMk4 + FP_SVMk4)

reca_SVMk4 <- (TP_SVMk4) / (TP_SVMk4 + FN_SVMk4)

f1_SVMk4 <- 2 * ((precision_SVMk4 * reca_SVMk4) / (precision_SVMk4 + reca_SVMk4))

cat("Accuracy of Titanic SVM:", acc_SVMk4, "\n")
cat("Precision of Titanic SVM:", precision_SVMk4, "\n")
cat("Recall of Titanic SVM:", reca_SVMk4, "\n")
cat("F1-Score of Titanic SVM:", f1_SVMk4, "\n")
```



# Naive Bayes(NB)
## NB with out cross-validation 

```{r}
library(e1071)

process_test$Sex <- factor(process_test$Sex, levels = levels(process_train$Sex))
process_test$Nickname <- factor(process_test$Nickname, levels = levels(process_train$Nickname))

set.seed(1)

naibawithoutcross <- naiveBayes(Survived ~ Pclass + Sex + Age + SibSp + Parch + Nickname, 
                             data = process_train)


summary(naibawithoutcross)
print(naibawithoutcross)


pred_naibawithoutcross <- predict(naibawithoutcross, newdata = process_test)


conmatrix_naibawithoutcross <- table(pred_naibawithoutcross, changedtpye_gender_submission$Survived)

print(conmatrix_naibawithoutcross)
```

```{r}
TP_naibawithoutcross <- sum(pred_naibawithoutcross == 1 & changedtpye_gender_submission$Survived == 1)

TN_naibawithoutcross <- sum(pred_naibawithoutcross == 0 & changedtpye_gender_submission$Survived == 0)

FP_naibawithoutcross <- sum(pred_naibawithoutcross == 1 & changedtpye_gender_submission$Survived == 0)

FN_naibawithoutcross <- sum(pred_naibawithoutcross == 0 & changedtpye_gender_submission$Survived == 1)

cat("True Positives in Titanic NB:", TP_naibawithoutcross, "\n")
cat("True Negatives in Titanic NB:", TN_naibawithoutcross, "\n")
cat("False Positives in Titanic NB:", FP_naibawithoutcross, "\n")
cat("False Negatives in Titanic NB:", FN_naibawithoutcross, "\n")
```

```{r}
table(pred_naibawithoutcross, changedtpye_gender_submission$Survived)

acc_naibawithoutcross <- (TP_naibawithoutcross + TN_naibawithoutcross) / (TP_naibawithoutcross + TN_naibawithoutcross + FP_naibawithoutcross + FN_naibawithoutcross)

precision_naibawithoutcross <- (TP_naibawithoutcross) / (TP_naibawithoutcross + FP_naibawithoutcross)

reca_naibawithoutcross <- (TP_naibawithoutcross) / (TP_naibawithoutcross + FN_naibawithoutcross)

f1_naibawithoutcross <- 2 * ((precision_naibawithoutcross * reca_naibawithoutcross) / (precision_naibawithoutcross + reca_naibawithoutcross))

cat("Accuracy of Titanic NB:", acc_naibawithoutcross, "\n")
cat("Precision of Titanic NB:", precision_naibawithoutcross, "\n")
cat("Recall of Titanic NB:", reca_naibawithoutcross, "\n")
cat("F1-Score of Titanic NB:", f1_naibawithoutcross, "\n")
```

## NB with k-fold cross-validation (k=10)

```{r}
unique(process_train$Nickname)
```

```{r}
unique(process_test$Nickname)
```

```{r}
unique(changedtpye_gender_submission$Nickname)
```


```{r}
library(caret)
library(dplyr)
library(forcats)


origprocess_train <- process_train
origprocess_test <- process_test

#Groups all Nickname levels that appear fewer than 5 times into an (Other) category
temprocess_train <- process_train %>%
  mutate(Nickname = fct_lump_min(Nickname, min = 5))

temprocess_test <- process_test %>%
  mutate(Nickname = fct_lump_min(Nickname, min = 5)) %>%
  mutate(Nickname = factor(Nickname, levels = levels(temprocess_train$Nickname)))

set.seed(1)
k10 <- trainControl(method = "cv", number = 10)

naibak10 <- train(Survived ~ Pclass + Sex + Age + SibSp + Parch + Nickname, 
                  data = temprocess_train, 
                  method = "nb", 
                  trControl = k10)

summary(naibak10)
print(naibak10)

pred_naibak10 <- predict(naibak10, newdata = temprocess_test)

conmatrix_naibak10 <- table(pred_naibak10, changedtpye_gender_submission$Survived)
print(conmatrix_naibak10)

process_train <- origprocess_train
process_test <- origprocess_test
```

```{r}
TP_naibak10 <- sum(pred_naibak10 == 1 & changedtpye_gender_submission$Survived == 1)

TN_naibak10 <- sum(pred_naibak10 == 0 & changedtpye_gender_submission$Survived == 0)

FP_naibak10 <- sum(pred_naibak10 == 1 & changedtpye_gender_submission$Survived == 0)

FN_naibak10 <- sum(pred_naibak10 == 0 & changedtpye_gender_submission$Survived == 1)

cat("True Positives in Titanic NB:", TP_naibak10, "\n")
cat("True Negatives in Titanic NB:", TN_naibak10, "\n")
cat("False Positives in Titanic NB:", FP_naibak10, "\n")
cat("False Negatives in Titanic NB:", FN_naibak10, "\n")
```

```{r}
table(pred_naibak10, changedtpye_gender_submission$Survived)

acc_naibak10 <- (TP_naibak10 + TN_naibak10) / (TP_naibak10 + TN_naibak10 + FP_naibak10 + FN_naibak10)

precision_naibak10 <- (TP_naibak10) / (TP_naibak10 + FP_naibak10)

reca_naibak10 <- (TP_naibak10) / (TP_naibak10 + FN_naibak10)

f1_naibak10 <- 2 * ((precision_naibak10 * reca_naibak10) / (precision_naibak10 + reca_naibak10))

cat("Accuracy of Titanic NB:", acc_naibak10, "\n")
cat("Precision of Titanic NB:", precision_naibak10, "\n")
cat("Recall of Titanic NB:", reca_naibak10, "\n")
cat("F1-Score of Titanic NB:", f1_naibak10, "\n")
```

## NB with k-fold cross-validation (k=5)

```{r}
library(caret)
library(dplyr)
library(forcats)


origprocess_train <- process_train
origprocess_test <- process_test

#Groups all Nickname levels that appear fewer than 5 times into an (Other) category
temprocess_train <- process_train %>%
  mutate(Nickname = fct_lump_min(Nickname, min = 5))

temprocess_test <- process_test %>%
  mutate(Nickname = fct_lump_min(Nickname, min = 5)) %>%
  mutate(Nickname = factor(Nickname, levels = levels(temprocess_train$Nickname)))

set.seed(1)
k5 <- trainControl(method = "cv", number = 5)

naibak5 <- train(Survived ~ Pclass + Sex + Age + SibSp + Parch + Nickname, 
                  data = temprocess_train, 
                  method = "nb", 
                  trControl = k5)

summary(naibak5)
print(naibak5)

pred_naibak5 <- predict(naibak5, newdata = temprocess_test)

conmatrix_naibak5 <- table(pred_naibak5, changedtpye_gender_submission$Survived)
print(conmatrix_naibak5)

process_train <- origprocess_train
process_test <- origprocess_test
```

```{r}
TP_naibak5 <- sum(pred_naibak5 == 1 & changedtpye_gender_submission$Survived == 1)

TN_naibak5 <- sum(pred_naibak5 == 0 & changedtpye_gender_submission$Survived == 0)

FP_naibak5 <- sum(pred_naibak5 == 1 & changedtpye_gender_submission$Survived == 0)

FN_naibak5 <- sum(pred_naibak5 == 0 & changedtpye_gender_submission$Survived == 1)

cat("True Positives in Titanic NB:", TP_naibak5, "\n")
cat("True Negatives in Titanic NB:", TN_naibak5, "\n")
cat("False Positives in Titanic NB:", FP_naibak5, "\n")
cat("False Negatives in Titanic NB:", FN_naibak5, "\n")
```

```{r}
table(pred_naibak5, changedtpye_gender_submission$Survived)

acc_naibak5 <- (TP_naibak5 + TN_naibak5) / (TP_naibak5 + TN_naibak5 + FP_naibak5 + FN_naibak5)

precision_naibak5 <- (TP_naibak5) / (TP_naibak5 + FP_naibak5)

reca_naibak5 <- (TP_naibak5) / (TP_naibak5 + FN_naibak5)

f1_naibak5 <- 2 * ((precision_naibak5 * reca_naibak5) / (precision_naibak5 + reca_naibak5))

cat("Accuracy of Titanic NB:", acc_naibak5, "\n")
cat("Precision of Titanic NB:", precision_naibak5, "\n")
cat("Recall of Titanic NB:", reca_naibak5, "\n")
cat("F1-Score of Titanic NB:", f1_naibak5, "\n")
```

## NB with k-fold cross-validation (k=4)

```{r}
library(caret)
library(dplyr)
library(forcats)


origprocess_train <- process_train
origprocess_test <- process_test

#Groups all Nickname levels that appear fewer than 5 times into an (Other) category
temprocess_train <- process_train %>%
  mutate(Nickname = fct_lump_min(Nickname, min = 5))

temprocess_test <- process_test %>%
  mutate(Nickname = fct_lump_min(Nickname, min = 5)) %>%
  mutate(Nickname = factor(Nickname, levels = levels(temprocess_train$Nickname)))

set.seed(1)
k4 <- trainControl(method = "cv", number = 4)

naibak4 <- train(Survived ~ Pclass + Sex + Age + SibSp + Parch + Nickname, 
                  data = temprocess_train, 
                  method = "nb", 
                  trControl = k4)

summary(naibak4)
print(naibak4)

pred_naibak4 <- predict(naibak4, newdata = temprocess_test)

conmatrix_naibak4 <- table(pred_naibak4, changedtpye_gender_submission$Survived)
print(conmatrix_naibak4)

process_train <- origprocess_train
process_test <- origprocess_test
```

```{r}
TP_naibak4 <- sum(pred_naibak4 == 1 & changedtpye_gender_submission$Survived == 1)

TN_naibak4 <- sum(pred_naibak4 == 0 & changedtpye_gender_submission$Survived == 0)

FP_naibak4 <- sum(pred_naibak4 == 1 & changedtpye_gender_submission$Survived == 0)

FN_naibak4 <- sum(pred_naibak4 == 0 & changedtpye_gender_submission$Survived == 1)

cat("True Positives in Titanic NB:", TP_naibak4, "\n")
cat("True Negatives in Titanic NB:", TN_naibak4, "\n")
cat("False Positives in Titanic NB:", FP_naibak4, "\n")
cat("False Negatives in Titanic NB:", FN_naibak4, "\n")
```

```{r}
table(pred_naibak4, changedtpye_gender_submission$Survived)

acc_naibak4 <- (TP_naibak4 + TN_naibak4) / (TP_naibak4 + TN_naibak4 + FP_naibak4 + FN_naibak4)

precision_naibak4 <- (TP_naibak4) / (TP_naibak4 + FP_naibak4)

reca_naibak4 <- (TP_naibak4) / (TP_naibak4 + FN_naibak4)

f1_naibak4 <- 2 * ((precision_naibak4 * reca_naibak4) / (precision_naibak4 + reca_naibak4))

cat("Accuracy of Titanic NB:", acc_naibak4, "\n")
cat("Precision of Titanic NB:", precision_naibak4, "\n")
cat("Recall of Titanic NB:", reca_naibak4, "\n")
cat("F1-Score of Titanic NB:", f1_naibak4, "\n")
```

### The best performance of Naive Bayes is without cross validuation, and it is higher a lot then with cross validuation, it is the same with article. 
### In my code when I use different number of K fold, Performance metrics will not change, it not the same with article, the article when k increase, Performance metrics increase a little bit.



# Decision Tree (DT)
## DT with out cross-validation 

```{r}
#install.packages("rpart")
```


```{r}
library(rpart)
library(caret)
library(dplyr)
library(forcats)

origprocess_train <- process_train
origprocess_test <- process_test

# Group all Nickname levels that appear fewer than 5 times into an (Other) category
temprocess_train <- process_train %>%
  mutate(Nickname = fct_lump_min(Nickname, min = 5))

temprocess_test <- process_test %>%
  mutate(Nickname = fct_lump_min(Nickname, min = 5)) %>%
  mutate(Nickname = factor(Nickname, levels = levels(temprocess_train$Nickname)))

set.seed(1)


treebestofdepth <- NULL
treebestofaccuracy <- 0

# try different depth of the tree
for(depth in c(5, 10, 20, 25, 30)) {
  set.seed(1)
  
  dectrwithoutcross <- rpart(Survived ~ Pclass + Sex + Age + SibSp + Parch + Nickname,
                             data = temprocess_train,
                             control = rpart.control(maxdepth = depth))

  print(paste("DT summary of max depth:", depth))
  summary(dectrwithoutcross)
  print(dectrwithoutcross)

  pred_dectrwithoutcross <- predict(dectrwithoutcross, newdata = temprocess_test, type = "class")

  conmatrix_dectrwithoutcross <- table(pred_dectrwithoutcross, changedtpye_gender_submission$Survived)
  print(conmatrix_dectrwithoutcross)

  accuracy <- sum(diag(conmatrix_dectrwithoutcross)) / sum(conmatrix_dectrwithoutcross)

  if(accuracy > treebestofaccuracy) {
    treebestofaccuracy <- accuracy
    treebestofdepth <- depth
  }
}
print(paste("Best Max Depth:", treebestofdepth, "with accuracy:", round(treebestofaccuracy * 100, 4), "%"))
process_train <- origprocess_train
process_test <- origprocess_test
```

```{r}
TP_dectrwithoutcross <- sum(pred_dectrwithoutcross == 1 & changedtpye_gender_submission$Survived == 1)

TN_dectrwithoutcross <- sum(pred_dectrwithoutcross == 0 & changedtpye_gender_submission$Survived == 0)

FP_dectrwithoutcross <- sum(pred_dectrwithoutcross == 1 & changedtpye_gender_submission$Survived == 0)

FN_dectrwithoutcross <- sum(pred_dectrwithoutcross == 0 & changedtpye_gender_submission$Survived == 1)

cat("True Positives in Titanic DT:", TP_dectrwithoutcross, "\n")
cat("True Negatives in Titanic DT:", TN_dectrwithoutcross, "\n")
cat("False Positives in Titanic DT:", FP_dectrwithoutcross, "\n")
cat("False Negatives in Titanic DT:", FN_dectrwithoutcross, "\n")
```

```{r}
# Check lengths
length(pred_dectrwithoutcross)
length(changedtpye_gender_submission$Survived)
```

```{r}
table(pred_dectrwithoutcross, changedtpye_gender_submission$Survived)

acc_dectrwithoutcross <- (TP_dectrwithoutcross + TN_dectrwithoutcross) / (TP_dectrwithoutcross + TN_dectrwithoutcross + FP_dectrwithoutcross + FN_dectrwithoutcross)

precision_dectrwithoutcross <- (TP_dectrwithoutcross) / (TP_dectrwithoutcross + FP_dectrwithoutcross)

reca_dectrwithoutcross <- (TP_dectrwithoutcross) / (TP_dectrwithoutcross + FN_dectrwithoutcross)

f1_dectrwithoutcross <- 2 * ((precision_dectrwithoutcross * reca_dectrwithoutcross) / (precision_dectrwithoutcross + reca_dectrwithoutcross))

cat("Accuracy of Titanic DT:", acc_dectrwithoutcross, "\n")
cat("Precision of Titanic DT:", precision_dectrwithoutcross, "\n")
cat("Recall of Titanic DT:", reca_dectrwithoutcross, "\n")
cat("F1-Score of Titanic DT:", f1_dectrwithoutcross, "\n")
```

## DT with k-fold cross-validation (k=10)

```{r}
library(rpart)
library(caret)
library(dplyr)
library(forcats)

origprocess_train <- process_train
origprocess_test <- process_test

# Group all Nickname levels that appear fewer than 5 times into an (Other) category
temprocess_train <- process_train %>%
  mutate(Nickname = fct_lump_min(Nickname, min = 5))

temprocess_test <- process_test %>%
  mutate(Nickname = fct_lump_min(Nickname, min = 5)) %>%
  mutate(Nickname = factor(Nickname, levels = levels(temprocess_train$Nickname)))

set.seed(1)


treebestofdepth <- NULL
treebestofaccuracy <- 0

k10 <- trainControl(method = "cv", number = 10)

# try different depth of the tree
sampledepth <- expand.grid(maxdepth = c(5, 10, 20, 25, 30))

dectrk10 <- train(Survived ~ Pclass + Sex + Age + SibSp + Parch + Nickname,
                  data = temprocess_train,
                  method = "rpart2",  
                  tuneGrid = sampledepth,
                  trControl = k10)

print(paste("DT summary of max depth:", depth))
summary(dectrk10)
print(dectrk10)


pred_dectrk10 <- predict(dectrk10, newdata = temprocess_test)

conmatrix_dectrk10 <- table(pred_dectrk10, changedtpye_gender_submission$Survived)
print(conmatrix_dectrk10)


treebestmodel <- dectrk10$bestTune


accuracy <- sum(diag(conmatrix_dectrk10)) / sum(conmatrix_dectrk10)

print(paste("Best Max Depth:", treebestmodel$maxdepth, "with accuracy:", round(accuracy * 100, 4), "%"))

process_train <- origprocess_train
process_test <- origprocess_test
```

```{r}
TP_dectrk10 <- sum(pred_dectrk10 == 1 & changedtpye_gender_submission$Survived == 1)

TN_dectrk10 <- sum(pred_dectrk10 == 0 & changedtpye_gender_submission$Survived == 0)

FP_dectrk10 <- sum(pred_dectrk10 == 1 & changedtpye_gender_submission$Survived == 0)

FN_dectrk10 <- sum(pred_dectrk10 == 0 & changedtpye_gender_submission$Survived == 1)

cat("True Positives in Titanic DT:", TP_dectrk10, "\n")
cat("True Negatives in Titanic DT:", TN_dectrk10, "\n")
cat("False Positives in Titanic DT:", FP_dectrk10, "\n")
cat("False Negatives in Titanic DT:", FN_dectrk10, "\n")
```

```{r}
table(pred_dectrk10, changedtpye_gender_submission$Survived)

acc_dectrk10 <- (TP_dectrk10 + TN_dectrk10) / (TP_dectrk10 + TN_dectrk10 + FP_dectrk10 + FN_dectrk10)

precision_dectrk10 <- (TP_dectrk10) / (TP_dectrk10 + FP_dectrk10)

reca_dectrk10 <- (TP_dectrk10) / (TP_dectrk10 + FN_dectrk10)

f1_dectrk10 <- 2 * ((precision_dectrk10 * reca_dectrk10) / (precision_dectrk10 + reca_dectrk10))

cat("Accuracy of Titanic DT:", acc_dectrk10, "\n")
cat("Precision of Titanic DT:", precision_dectrk10, "\n")
cat("Recall of Titanic DT:", reca_dectrk10, "\n")
cat("F1-Score of Titanic DT:", f1_dectrk10, "\n")
```



## DT with k-fold cross-validation (k=5)

```{r}
library(rpart)
library(caret)
library(dplyr)
library(forcats)

origprocess_train <- process_train
origprocess_test <- process_test

# Group all Nickname levels that appear fewer than 5 times into an (Other) category
temprocess_train <- process_train %>%
  mutate(Nickname = fct_lump_min(Nickname, min = 5))

temprocess_test <- process_test %>%
  mutate(Nickname = fct_lump_min(Nickname, min = 5)) %>%
  mutate(Nickname = factor(Nickname, levels = levels(temprocess_train$Nickname)))

set.seed(1)


treebestofdepth <- NULL
treebestofaccuracy <- 0

k5 <- trainControl(method = "cv", number = 5)

# try different depth of the tree
sampledepth <- expand.grid(maxdepth = c(5, 10, 20, 25, 30))

dectrk5 <- train(Survived ~ Pclass + Sex + Age + SibSp + Parch + Nickname,
                  data = temprocess_train,
                  method = "rpart2",  
                  tuneGrid = sampledepth,
                  trControl = k5)

print(paste("DT summary of max depth:", depth))
summary(dectrk5)
print(dectrk5)


pred_dectrk5 <- predict(dectrk5, newdata = temprocess_test)

conmatrix_dectrk5 <- table(pred_dectrk5, changedtpye_gender_submission$Survived)
print(conmatrix_dectrk5)


treebestmodel <- dectrk5$bestTune


accuracy <- sum(diag(conmatrix_dectrk5)) / sum(conmatrix_dectrk5)

print(paste("Best Max Depth:", treebestmodel$maxdepth, "with accuracy:", round(accuracy * 100, 4), "%"))

process_train <- origprocess_train
process_test <- origprocess_test
```

```{r}
TP_dectrk5 <- sum(pred_dectrk5 == 1 & changedtpye_gender_submission$Survived == 1)

TN_dectrk5 <- sum(pred_dectrk5 == 0 & changedtpye_gender_submission$Survived == 0)

FP_dectrk5 <- sum(pred_dectrk5 == 1 & changedtpye_gender_submission$Survived == 0)

FN_dectrk5 <- sum(pred_dectrk5 == 0 & changedtpye_gender_submission$Survived == 1)

cat("True Positives in Titanic DT:", TP_dectrk5, "\n")
cat("True Negatives in Titanic DT:", TN_dectrk5, "\n")
cat("False Positives in Titanic DT:", FP_dectrk5, "\n")
cat("False Negatives in Titanic DT:", FN_dectrk5, "\n")
```

```{r}
table(pred_dectrk5, changedtpye_gender_submission$Survived)

acc_dectrk5 <- (TP_dectrk5 + TN_dectrk5) / (TP_dectrk5 + TN_dectrk5 + FP_dectrk5 + FN_dectrk5)

precision_dectrk5 <- (TP_dectrk5) / (TP_dectrk5 + FP_dectrk5)

reca_dectrk5 <- (TP_dectrk5) / (TP_dectrk5 + FN_dectrk5)

f1_dectrk5 <- 2 * ((precision_dectrk5 * reca_dectrk5) / (precision_dectrk5 + reca_dectrk5))

cat("Accuracy of Titanic DT:", acc_dectrk5, "\n")
cat("Precision of Titanic DT:", precision_dectrk5, "\n")
cat("Recall of Titanic DT:", reca_dectrk5, "\n")
cat("F1-Score of Titanic DT:", f1_dectrk5, "\n")
```



## DT with k-fold cross-validation (k=4)

```{r}
library(rpart)
library(caret)
library(dplyr)
library(forcats)

origprocess_train <- process_train
origprocess_test <- process_test

# Group all Nickname levels that appear fewer than 5 times into an (Other) category
temprocess_train <- process_train %>%
  mutate(Nickname = fct_lump_min(Nickname, min = 5))

temprocess_test <- process_test %>%
  mutate(Nickname = fct_lump_min(Nickname, min = 5)) %>%
  mutate(Nickname = factor(Nickname, levels = levels(temprocess_train$Nickname)))

set.seed(1)


treebestofdepth <- NULL
treebestofaccuracy <- 0

k4 <- trainControl(method = "cv", number = 4)

# try different depth of the tree
sampledepth <- expand.grid(maxdepth = c(5, 10, 20, 25, 30))

dectrk4 <- train(Survived ~ Pclass + Sex + Age + SibSp + Parch + Nickname,
                  data = temprocess_train,
                  method = "rpart2",  
                  tuneGrid = sampledepth,
                  trControl = k4)

print(paste("DT summary of max depth:", depth))
summary(dectrk4)
print(dectrk4)


pred_dectrk4 <- predict(dectrk4, newdata = temprocess_test)

conmatrix_dectrk4 <- table(pred_dectrk4, changedtpye_gender_submission$Survived)
print(conmatrix_dectrk4)


treebestmodel <- dectrk4$bestTune


accuracy <- sum(diag(conmatrix_dectrk4)) / sum(conmatrix_dectrk4)

print(paste("Best Max Depth:", treebestmodel$maxdepth, "with accuracy:", round(accuracy * 100, 4), "%"))

process_train <- origprocess_train
process_test <- origprocess_test
```

```{r}
TP_dectrk4 <- sum(pred_dectrk4 == 1 & changedtpye_gender_submission$Survived == 1)

TN_dectrk4 <- sum(pred_dectrk4 == 0 & changedtpye_gender_submission$Survived == 0)

FP_dectrk4 <- sum(pred_dectrk4 == 1 & changedtpye_gender_submission$Survived == 0)

FN_dectrk4 <- sum(pred_dectrk4 == 0 & changedtpye_gender_submission$Survived == 1)

cat("True Positives in Titanic DT:", TP_dectrk4, "\n")
cat("True Negatives in Titanic DT:", TN_dectrk4, "\n")
cat("False Positives in Titanic DT:", FP_dectrk4, "\n")
cat("False Negatives in Titanic DT:", FN_dectrk4, "\n")
```

```{r}
table(pred_dectrk4, changedtpye_gender_submission$Survived)

acc_dectrk4 <- (TP_dectrk4 + TN_dectrk4) / (TP_dectrk4 + TN_dectrk4 + FP_dectrk4 + FN_dectrk4)

precision_dectrk4 <- (TP_dectrk4) / (TP_dectrk4 + FP_dectrk4)

reca_dectrk4 <- (TP_dectrk4) / (TP_dectrk4 + FN_dectrk4)

f1_dectrk4 <- 2 * ((precision_dectrk4 * reca_dectrk4) / (precision_dectrk4 + reca_dectrk4))

cat("Accuracy of Titanic DT:", acc_dectrk4, "\n")
cat("Precision of Titanic DT:", precision_dectrk4, "\n")
cat("Recall of Titanic DT:", reca_dectrk4, "\n")
cat("F1-Score of Titanic DT:", f1_dectrk4, "\n")
```



# Random Forest(RF)
## RF with out cross-validation 

```{r}
library(randomForest)

process_test$Sex <- factor(process_test$Sex, levels = levels(process_train$Sex))
process_test$Nickname <- factor(process_test$Nickname, levels = levels(process_train$Nickname))

set.seed(1)

# find the best tree
samplenumtree <- c(10, 100, 250, 500, 1000)
samplemaxnodes <- c(10, 20, 30, 50, 100)

setbestnumtree <- 0
setbestmaxnodes <- 0
setbestaccuracy <- 0

for (ntree in samplenumtree) {
  for (maxnodes in samplemaxnodes) {
    ranfomodel <- randomForest(Survived ~ Pclass + Sex + Age + SibSp + Parch + Nickname, 
                             data = process_train, 
                             ntree = ntree,
                             maxnodes = maxnodes)
  
  pred_ranfomodel <- predict(ranfomodel, newdata = process_test)
  
  conmatrix <- table(pred_ranfomodel, changedtpye_gender_submission$Survived)
  
  accuracy_nonbest_ranfo <- sum(diag(conmatrix)) / sum(conmatrix)
  
  print(paste("Number of tree =", ntree, "num Max nodes =", maxnodes, "num Accuracy =", accuracy_nonbest_ranfo))
  
    if (accuracy_nonbest_ranfo > setbestaccuracy) {
      setbestaccuracy <- accuracy_nonbest_ranfo
      setbestnumtree <- ntree
      setbestmaxnodes <- maxnodes
    }
  }
}

print(paste("Best number tree in random forest model =", setbestnumtree, 
            "and the best max nodes =", setbestmaxnodes, 
            "with the best accuracy =", setbestaccuracy))

# use best n tree to do model
ranfowithoutcross <- randomForest(Survived ~ Pclass + Sex + Age + SibSp + Parch + Nickname, 
                                data = process_train, 
                                ntree = setbestnumtree, 
                                maxnodes = setbestmaxnodes)

pred_ranfowithoutcross <- predict(ranfowithoutcross, newdata = process_test)

conmatrix_ranfowithoutcross <- table(pred_ranfowithoutcross, changedtpye_gender_submission$Survived)

print(conmatrix_ranfowithoutcross)
```


```{r}
summary(ranfomodel)
print(ranfomodel)
```

```{r}
summary(ranfowithoutcross)
print(ranfowithoutcross)
```


```{r}
TP_ranfowithoutcross <- sum(pred_ranfowithoutcross == 1 & changedtpye_gender_submission$Survived == 1)

TN_ranfowithoutcross <- sum(pred_ranfowithoutcross == 0 & changedtpye_gender_submission$Survived == 0)

FP_ranfowithoutcross <- sum(pred_ranfowithoutcross == 1 & changedtpye_gender_submission$Survived == 0)

FN_ranfowithoutcross <- sum(pred_ranfowithoutcross == 0 & changedtpye_gender_submission$Survived == 1)

cat("True Positives in Titanic RF:", TP_ranfowithoutcross, "\n")
cat("True Negatives in Titanic RF:", TN_ranfowithoutcross, "\n")
cat("False Positives in Titanic RF:", FP_ranfowithoutcross, "\n")
cat("False Negatives in Titanic RF:", FN_ranfowithoutcross, "\n")
```

```{r}
table(pred_ranfowithoutcross, changedtpye_gender_submission$Survived)

acc_ranfowithoutcross <- (TP_ranfowithoutcross + TN_ranfowithoutcross) / (TP_ranfowithoutcross + TN_ranfowithoutcross + FP_ranfowithoutcross + FN_ranfowithoutcross)

precision_ranfowithoutcross <- (TP_ranfowithoutcross) / (TP_ranfowithoutcross + FP_ranfowithoutcross)

reca_ranfowithoutcross <- (TP_ranfowithoutcross) / (TP_ranfowithoutcross + FN_ranfowithoutcross)

f1_ranfowithoutcross <- 2 * ((precision_ranfowithoutcross * reca_ranfowithoutcross) / (precision_ranfowithoutcross + reca_ranfowithoutcross))

cat("Accuracy of Titanic RF:", acc_ranfowithoutcross, "\n")
cat("Precision of Titanic RF:", precision_ranfowithoutcross, "\n")
cat("Recall of Titanic RF:", reca_ranfowithoutcross, "\n")
cat("F1-Score of Titanic RF:", f1_ranfowithoutcross, "\n")
```



## RF with k-fold cross-validation (k=10)

```{r}
library(randomForest)

process_test$Sex <- factor(process_test$Sex, levels = levels(process_train$Sex))
process_test$Nickname <- factor(process_test$Nickname, levels = levels(process_train$Nickname))

set.seed(1)

k10 <- trainControl(method = "cv", number = 10)

#ranging from 1 to the number of predictors minus one
creatgrid <- expand.grid(mtry = 1:(ncol(process_train) - 1))  

# find the best tree
samplenumtree <- c(10, 100, 250, 500, 1000)
samplemaxnodes <- c(10, 20, 30, 50, 100)

setbestnumtree <- 0
setbestmaxnodes <- 0
setbestaccuracy <- 0

for (ntree in samplenumtree) {
  for (maxnodes in samplemaxnodes) {
    ranfomodel <- train(Survived ~ Pclass + Sex + Age + SibSp + Parch + Nickname, 
                    data = process_train,
                    method = "rf",
                    tuneGrid = creatgrid, 
                    trControl = k10,       
                    ntree = ntree,         
                    maxnodes = maxnodes)
  
  pred_ranfomodel <- predict(ranfomodel, newdata = process_test)
  
  conmatrix <- table(pred_ranfomodel, changedtpye_gender_submission$Survived)
  
  accuracy_nonbest_ranfo <- sum(diag(conmatrix)) / sum(conmatrix)
  
  print(paste("Number of tree =", ntree, "num Max nodes =", maxnodes, "num Accuracy =", accuracy_nonbest_ranfo))
  
    if (accuracy_nonbest_ranfo > setbestaccuracy) {
      setbestaccuracy <- accuracy_nonbest_ranfo
      setbestnumtree <- ntree
      setbestmaxnodes <- maxnodes
    }
  }
}

print(paste("Best number tree in random forest model =", setbestnumtree, 
            "and the best max nodes =", setbestmaxnodes, 
            "with the best accuracy =", setbestaccuracy))

# use best n tree to do model
k10 <- trainControl(method = "cv", number = 10)
ranfok10 <- train(Survived ~ Pclass + Sex + Age + SibSp + Parch + Nickname, 
                                data = process_train,
                                method = "rf",
                                tuneGrid = creatgrid,
                                trControl = k10,
                                ntree = setbestnumtree, 
                                maxnodes = setbestmaxnodes
                                )

pred_ranfok10 <- predict(ranfok10, newdata = process_test)

conmatrix_ranfok10 <- table(pred_ranfok10, changedtpye_gender_submission$Survived)

print(conmatrix_ranfok10)
```

```{r}
summary(ranfomodel)
print(ranfomodel)
```

```{r}
summary(ranfok10)
print(ranfok10)
```

```{r}
TP_ranfok10 <- sum(pred_ranfok10 == 1 & changedtpye_gender_submission$Survived == 1)

TN_ranfok10 <- sum(pred_ranfok10 == 0 & changedtpye_gender_submission$Survived == 0)

FP_ranfok10 <- sum(pred_ranfok10 == 1 & changedtpye_gender_submission$Survived == 0)

FN_ranfok10 <- sum(pred_ranfok10 == 0 & changedtpye_gender_submission$Survived == 1)

cat("True Positives in Titanic RF:", TP_ranfok10, "\n")
cat("True Negatives in Titanic RF:", TN_ranfok10, "\n")
cat("False Positives in Titanic RF:", FP_ranfok10, "\n")
cat("False Negatives in Titanic RF:", FN_ranfok10, "\n")
```

```{r}
table(pred_ranfok10, changedtpye_gender_submission$Survived)

acc_ranfok10 <- (TP_ranfok10 + TN_ranfok10) / (TP_ranfok10 + TN_ranfok10 + FP_ranfok10 + FN_ranfok10)

precision_ranfok10 <- (TP_ranfok10) / (TP_ranfok10 + FP_ranfok10)

reca_ranfok10 <- (TP_ranfok10) / (TP_ranfok10 + FN_ranfok10)

f1_ranfok10 <- 2 * ((precision_ranfok10 * reca_ranfok10) / (precision_ranfok10 + reca_ranfok10))

cat("Accuracy of Titanic RF:", acc_ranfok10, "\n")
cat("Precision of Titanic RF:", precision_ranfok10, "\n")
cat("Recall of Titanic RF:", reca_ranfok10, "\n")
cat("F1-Score of Titanic RF:", f1_ranfok10, "\n")
```



## RF with k-fold cross-validation (k=5)

```{r}
library(randomForest)

process_test$Sex <- factor(process_test$Sex, levels = levels(process_train$Sex))
process_test$Nickname <- factor(process_test$Nickname, levels = levels(process_train$Nickname))

set.seed(1)

k5 <- trainControl(method = "cv", number = 5)

#ranging from 1 to the number of predictors minus one
creatgrid <- expand.grid(mtry = 1:(ncol(process_train) - 1))  

# find the best tree
samplenumtree <- c(10, 100, 250, 500, 1000)
samplemaxnodes <- c(10, 20, 30, 50, 100)

setbestnumtree <- 0
setbestmaxnodes <- 0
setbestaccuracy <- 0

for (ntree in samplenumtree) {
  for (maxnodes in samplemaxnodes) {
    ranfomodel <- train(Survived ~ Pclass + Sex + Age + SibSp + Parch + Nickname, 
                    data = process_train,
                    method = "rf",
                    tuneGrid = creatgrid, 
                    trControl = k5,       
                    ntree = ntree,         
                    maxnodes = maxnodes)
  
  pred_ranfomodel <- predict(ranfomodel, newdata = process_test)
  
  conmatrix <- table(pred_ranfomodel, changedtpye_gender_submission$Survived)
  
  accuracy_nonbest_ranfo <- sum(diag(conmatrix)) / sum(conmatrix)
  
  print(paste("Number of tree =", ntree, "num Max nodes =", maxnodes, "num Accuracy =", accuracy_nonbest_ranfo))
  
    if (accuracy_nonbest_ranfo > setbestaccuracy) {
      setbestaccuracy <- accuracy_nonbest_ranfo
      setbestnumtree <- ntree
      setbestmaxnodes <- maxnodes
    }
  }
}

print(paste("Best number tree in random forest model =", setbestnumtree, 
            "and the best max nodes =", setbestmaxnodes, 
            "with the best accuracy =", setbestaccuracy))

# use best n tree to do model
k5 <- trainControl(method = "cv", number = 5)
ranfok5 <- train(Survived ~ Pclass + Sex + Age + SibSp + Parch + Nickname, 
                                data = process_train,
                                method = "rf",
                                tuneGrid = creatgrid,
                                trControl = k5,
                                ntree = setbestnumtree, 
                                maxnodes = setbestmaxnodes
                                )

pred_ranfok5 <- predict(ranfok5, newdata = process_test)

conmatrix_ranfok5 <- table(pred_ranfok5, changedtpye_gender_submission$Survived)

print(conmatrix_ranfok5)
```

```{r}
summary(ranfomodel)
print(ranfomodel)
```

```{r}
summary(ranfok5)
print(ranfok5)
```

```{r}
TP_ranfok5 <- sum(pred_ranfok5 == 1 & changedtpye_gender_submission$Survived == 1)

TN_ranfok5 <- sum(pred_ranfok5 == 0 & changedtpye_gender_submission$Survived == 0)

FP_ranfok5 <- sum(pred_ranfok5 == 1 & changedtpye_gender_submission$Survived == 0)

FN_ranfok5 <- sum(pred_ranfok5 == 0 & changedtpye_gender_submission$Survived == 1)

cat("True Positives in Titanic RF:", TP_ranfok5, "\n")
cat("True Negatives in Titanic RF:", TN_ranfok5, "\n")
cat("False Positives in Titanic RF:", FP_ranfok5, "\n")
cat("False Negatives in Titanic RF:", FN_ranfok5, "\n")
```

```{r}
table(pred_ranfok5, changedtpye_gender_submission$Survived)

acc_ranfok5 <- (TP_ranfok5 + TN_ranfok5) / (TP_ranfok5 + TN_ranfok5 + FP_ranfok5 + FN_ranfok5)

precision_ranfok5 <- (TP_ranfok5) / (TP_ranfok5 + FP_ranfok5)

reca_ranfok5 <- (TP_ranfok5) / (TP_ranfok5 + FN_ranfok5)

f1_ranfok5 <- 2 * ((precision_ranfok5 * reca_ranfok5) / (precision_ranfok5 + reca_ranfok5))

cat("Accuracy of Titanic RF:", acc_ranfok5, "\n")
cat("Precision of Titanic RF:", precision_ranfok5, "\n")
cat("Recall of Titanic RF:", reca_ranfok5, "\n")
cat("F1-Score of Titanic RF:", f1_ranfok5, "\n")
```



## RF with k-fold cross-validation (k=4)

```{r}
library(randomForest)

process_test$Sex <- factor(process_test$Sex, levels = levels(process_train$Sex))
process_test$Nickname <- factor(process_test$Nickname, levels = levels(process_train$Nickname))

set.seed(1)

k4 <- trainControl(method = "cv", number = 4)

#ranging from 1 to the number of predictors minus one
creatgrid <- expand.grid(mtry = 1:(ncol(process_train) - 1))  

# find the best tree
samplenumtree <- c(10, 100, 250, 500, 1000)
samplemaxnodes <- c(10, 20, 30, 50, 100)

setbestnumtree <- 0
setbestmaxnodes <- 0
setbestaccuracy <- 0

for (ntree in samplenumtree) {
  for (maxnodes in samplemaxnodes) {
    ranfomodel <- train(Survived ~ Pclass + Sex + Age + SibSp + Parch + Nickname, 
                    data = process_train,
                    method = "rf",
                    tuneGrid = creatgrid, 
                    trControl = k4,       
                    ntree = ntree,         
                    maxnodes = maxnodes)
  
  pred_ranfomodel <- predict(ranfomodel, newdata = process_test)
  
  conmatrix <- table(pred_ranfomodel, changedtpye_gender_submission$Survived)
  
  accuracy_nonbest_ranfo <- sum(diag(conmatrix)) / sum(conmatrix)
  
  print(paste("Number of tree =", ntree, "num Max nodes =", maxnodes, "num Accuracy =", accuracy_nonbest_ranfo))
  
    if (accuracy_nonbest_ranfo > setbestaccuracy) {
      setbestaccuracy <- accuracy_nonbest_ranfo
      setbestnumtree <- ntree
      setbestmaxnodes <- maxnodes
    }
  }
}

print(paste("Best number tree in random forest model =", setbestnumtree, 
            "and the best max nodes =", setbestmaxnodes, 
            "with the best accuracy =", setbestaccuracy))

# use best n tree to do model
k4 <- trainControl(method = "cv", number = 4)
ranfok4 <- train(Survived ~ Pclass + Sex + Age + SibSp + Parch + Nickname, 
                                data = process_train,
                                method = "rf",
                                tuneGrid = creatgrid,
                                trControl = k4,
                                ntree = setbestnumtree, 
                                maxnodes = setbestmaxnodes
                                )

pred_ranfok4 <- predict(ranfok4, newdata = process_test)

conmatrix_ranfok4 <- table(pred_ranfok4, changedtpye_gender_submission$Survived)

print(conmatrix_ranfok4)
```

```{r}
summary(ranfomodel)
print(ranfomodel)
```

```{r}
summary(ranfok4)
print(ranfok4)
```

```{r}
TP_ranfok4 <- sum(pred_ranfok4 == 1 & changedtpye_gender_submission$Survived == 1)

TN_ranfok4 <- sum(pred_ranfok4 == 0 & changedtpye_gender_submission$Survived == 0)

FP_ranfok4 <- sum(pred_ranfok4 == 1 & changedtpye_gender_submission$Survived == 0)

FN_ranfok4 <- sum(pred_ranfok4 == 0 & changedtpye_gender_submission$Survived == 1)

cat("True Positives in Titanic RF:", TP_ranfok4, "\n")
cat("True Negatives in Titanic RF:", TN_ranfok4, "\n")
cat("False Positives in Titanic RF:", FP_ranfok4, "\n")
cat("False Negatives in Titanic RF:", FN_ranfok4, "\n")
```

```{r}
table(pred_ranfok4, changedtpye_gender_submission$Survived)

acc_ranfok4 <- (TP_ranfok4 + TN_ranfok4) / (TP_ranfok4 + TN_ranfok4 + FP_ranfok4 + FN_ranfok4)

precision_ranfok4 <- (TP_ranfok4) / (TP_ranfok4 + FP_ranfok4)

reca_ranfok4 <- (TP_ranfok4) / (TP_ranfok4 + FN_ranfok4)

f1_ranfok4 <- 2 * ((precision_ranfok4 * reca_ranfok4) / (precision_ranfok4 + reca_ranfok4))

cat("Accuracy of Titanic RF:", acc_ranfok4, "\n")
cat("Precision of Titanic RF:", precision_ranfok4, "\n")
cat("Recall of Titanic RF:", reca_ranfok4, "\n")
cat("F1-Score of Titanic RF:", f1_ranfok4, "\n")
```



















